Notes on scheduling the SAA2019 meeting

### 2017-11-06 analysis/paper/saa2018_posters_knapsack.Rmd
divide up poster sessions with >8 papers and schedule

We can combine general and organised sessions, for the purpose of scheduling. BUt we have to get the after hours sessions and put them aside.

@@

And we'll combine day-parts that are in the same paper day-part, which will make it simper for avoiding clashes.

```{r}
# exclude Thurs evening from now on
poster_session_durations_no_th_ev <- 
  poster_session_durations %>% 
  filter(day_part != "Th-evening") %>% 
  group_by(day_part) %>% 
  summarise(n_max = sum(`N max`) )

poster_session_durations_no_th_ev
```

We want to avoid people giving a poster and a talk at the same time, so we need to check with the paper schedule:

```{r}
sessions_in_day_parts_data_with_all_names_df <- 
  read_csv("../data/derived_data/sessions_in_day_parts_data_with_all_names_df.csv")
```

Now we need to put the poster sessions into the remaining slots while checking to see if one of the poster authors is already scheduled to give a paper.

```{r}
# to avoid overfilling day-parts
buffer <-7 # is the largest to avoid overfill

# day-parts to fill
 poster_session_durations_no_th_ev <- 
   poster_session_durations_no_th_ev %>% 
   arrange(desc(n_max))

# make a unique value for names for all posters
all_posters_individual_no_ah <- 
  all_posters_individual_no_ah %>% 
    mutate(uid = glue('{`First Name`}_{`Last Name`}'))

# get table of session id's only 
poster_session_ids <- 
unique(all_posters_individual_no_ah$`Session Id`)

# we're going to change these
  all_posters_individual_no_ah_tmp <- all_posters_individual_no_ah
  container_for_this_day_part <- 0
  poster_session_durations_no_th_ev$numbers_of_posters <- 0
  poster_session_durations_no_th_ev$n_posters <- 0
  poster_session_durations_no_th_ev$n_sessions <- 0
  poster_session_durations_no_th_ev$c_sessions <- 0
  poster_session_durations_no_th_ev$c_posters <- 0
  all_of_the_day_parts <- 0
  i <- 1

# loop over each day-slot available for poster sessions
# i is each day-part
for(i in seq_len(nrow(poster_session_durations_no_th_ev))) {

    # stop loop gracefully when we have assigned all people to day-parts
  if (nrow(all_posters_individual_no_ah_tmp) == 0){ break }

    # how many rows (posters) to we have remaning to allocate to a day-part?
  print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))

  # need a buffer to avoid overfilling day-parts
  target_capacity <- poster_session_durations_no_th_ev$n_max[i] - buffer

  # prepare variables for the next round, just to be sure
  cumulative_number_of_posters <- 0
  container_for_this_day_part <- 0
  this_session_posters <- 0
  # j is each poster session
  j <- 1

  while(cumulative_number_of_posters <= target_capacity & 
        j <= length(poster_session_ids)) {

  # get a session 
  this_session <- poster_session_ids[j]

  # get names for all the people in this poster session
  this_session_posters <- 
    all_posters_individual_no_ah_tmp %>% 
      filter(`Session Id` == this_session)

#-------------------------------------------------------------------  
  # check to see if these names are already in this day-part
  # or in the paper sessions
  if (j == 1){

    # do nothing is this is the first incoming session 
    names_already_in <-  0
    names_already_in2 <- 0

  } else {
      # get distinct names for each session previously added
      # to this day-part
      # (don't care about same name multiple times in one session)
      distinct_names_for_each_session_so_far <- 
      container_for_this_day_part %>% 
        group_by(uid) %>% 
        distinct(uid, .keep_all = TRUE) %>% 
        pull(uid)

      # check to see if incoming session has 
      # names that already exist in this day_part
      names_already_in <- sum(this_session_posters$uid %in% distinct_names_for_each_session_so_far)

      # check to see if papers for this day-part already have the name in it
      names_papers_in_this_day_part <- 
        sessions_in_day_parts_data_with_all_names_df %>% 
        filter(day_part_from_while.y == poster_session_durations_no_th_ev[i,]$day_part) %>% 
        pull(uid.y)

      names_already_in2 <- sum(this_session_posters$uid %in% names_papers_in_this_day_part)

  }

#-------------------------------------------------------------------
    # if there are no name conflicts, then
    # accumulate the incoming session onto the sessions
    # we've already added to this day-part

    if (j == 1) {

       container_for_this_day_part <- 
        this_session_posters

      } else {
      if (names_already_in == 0 & names_already_in2 == 0) {

      # if there are no name clashes, 
      # add these names to previous names in this day-part
      # bind rows of session that has no clashing  names with 
      # sessions we already added to this day-part
      container_for_this_day_part <- 
        bind_rows(this_session_posters,
                  container_for_this_day_part)

    }  else {

      # do not add this session to the day-part 
      # because at least one of the names is already in there
      }
      }

#-------------------------------------------------------------------  
  # compute the number of posters we've added to this day-part
  # by counting the abstract Id, because this excludes the session chairs
    cumulative_number_of_posters <- 
    nrow(container_for_this_day_part)

   # store this result
    poster_session_durations_no_th_ev$n_posters[i] <- cumulative_number_of_posters

    print(glue('cumulative number of posters: {cumulative_number_of_posters}, j = {j}, i = {i}'))

  # go to next session
  j <- j + 1
    }
  # end of while loop

   # add an identifier so we can see where we have assigned each session so far
 container_for_this_day_part_with_session_from_while <- 
   container_for_this_day_part %>% 
   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$day_part[i]))

 # bind rows to previous day-parts we might have done
  if (i == 1) {

  all_of_the_day_parts <- container_for_this_day_part_with_session_from_while

  } else {

    # check to see

  all_of_the_day_parts <- bind_rows(container_for_this_day_part_with_session_from_while, 
                                      all_of_the_day_parts)
  }

 poster_session_durations_no_th_ev$n_sessions[i] <- length(unique(container_for_this_day_part_with_session_from_while$`Session Id`))

 poster_session_durations_no_th_ev$c_sessions[i] <-  sum(poster_session_durations_no_th_ev$n_sessions)
 poster_session_durations_no_th_ev$c_posters[i] <-  sum(poster_session_durations_no_th_ev$n_posters)

 print(poster_session_durations_no_th_ev %>% select(n_max, 
                                                   day_part,
                                                   n_posters,
                                                   n_sessions, 
                                                   c_sessions,
                                                   c_posters))

 # for the next day-part, we need to make sure
 # we don't add the same session twice to different day-parts
 # so let's update our big table of posters, and drop the session that 
 # we've already got in the day-part

all_posters_individual_no_ah_tmp <-  
   all_posters_individual_no_ah_tmp %>% 
   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))

}
# end of for loop

@@

```{r}
# we have 532 posters not in ah
x1 <- 
all_posters_individual_no_ah

x2 <- 
all_of_the_day_parts

# what sessions are not in there?
session_output <- 
x2 %>% 
  group_by(`Session Id`) %>% 
  pull(`Session Id`) %>% 
  unique

session_input <- 
  x1 %>% 
  group_by(`Session Id`) %>% 
  pull(`Session Id`) %>% 
  unique

setdiff(session_input, session_output)

x1 %>% 
  filter(is.na(Title))

# what are in the output of this loop that is not in the input?
# duplicates?
all_of_the_day_parts %>% 
  distinct()

# why is each poster in each day-part?
all_of_the_day_parts %>% 
 arrange(uid) %>% View

# how many abstracts
posters_with_abstract_ids <- 
all_of_the_day_parts %>% 
  filter(!is.na(`Abstract Id`)) %>% 
  group_by(day_part_from_while) %>% 
  tally() %>% 
  summarise(sum(n)) %>% 
  pull()
```

We have `r posters_with_abstract_ids` posters with Abstract IDs (and so not chair/co-chairs). Compare to `r nrow(gen_and_org_poster_session_posters)` posters that we started with, and space available for `r sum(poster_session_durations$'N max')` posters.

>>>>

We can combine general and organised sessions, for the purpose of scheduling. But we have to get the after hours sessions and put them aside.

@@

```{r}
# see if we can do the poster day-parts individually

poster_session_durations_no_th_ev <- 
  poster_session_durations %>% 
  filter(day_part != "Th-evening") %>% 
  mutate(poster_day_part = glue('poster-{day_part}')) %>% 
  group_by(poster_day_part) %>% 
  mutate(label = 1:n()) %>% 
  ungroup() %>% 
  mutate(poster_day_part = glue('{poster_day_part}_{label}')) %>% 
  select(-label) %>% 
  mutate(n_max = `N max`)
```

We're not going to bother with clashes between people giving posters and people giving papers, because they are both 'presenting' roles, and a person can only have one of those.

```{r}
# check 
people_with_2_non_poster_roles <- 
all_abstracts %>% 
  group_by(`First Name`, `Last Name`) %>% 
  add_tally() %>% 
  filter(n >= 2) %>% 
  filter(!grepl("Poster", `Session Type`)) %>% 
  mutate(uid = glue('{`First Name`} {`Last Name`}'))

people_in_posters_with_2_roles <- 
  all_abstracts %>% 
  group_by(`First Name`, `Last Name`) %>% 
  add_tally() %>% 
  filter(n >= 2) %>% 
  filter(grepl("Poster", `Session Type`)) %>% 
  mutate(uid = glue('{`First Name`} {`Last Name`}'))

# any overlap?
people_with_posters_and_another_role <- 
people_in_posters_with_2_roles$uid[people_in_posters_with_2_roles$uid %in% people_with_2_non_poster_roles$uid]
```

So let's go right onto the knapsack to see if we can fit them all in.

```{r}

# number of posters in each session
posters_to_allocate <- 
  all_posters_individual_no_ah %>% 
  group_by(`Session Id`) %>% 
  tally() %>% 
  pull(n)
sum(posters_to_allocate)

# size of each slot to fill with posters
poster_sessions_available <- poster_session_durations_no_th_ev$n_max
sum(poster_session_durations_no_th_ev$n_max)

library(adagio)
# mknapsack calling signature is: mknapsack(values, weights, capacities)
solution <- mknapsack(posters_to_allocate, 
                      posters_to_allocate, 
                      poster_sessions_available)

solution$ksack

# poster sessions only
all_posters_individual_no_ah_sessions <- 
  all_posters_individual_no_ah %>% 
  group_by(`Session Id`) %>% 
  select(`Session Id`) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(ksack = solution$ksack)

# put the results back on indiv posters
all_posters_individual_no_ah_ksack <- 
  all_posters_individual_no_ah %>% 
  left_join(all_posters_individual_no_ah_sessions)

all_posters_individual_no_ah_ksack %>%

@@

We may need to split up some of the general poster sessions to make it easier to pack them in. What is the distribution of session sizes? What if we chopped in half all the sessions bigger than 12 posters?

We have no clash with paper presenters.

### 2017-11-06 analysis/paper/saa2018_posters_while_loop.Rmd
divide up poster sessions with >8 papers and schedule

# Here we get all the general session posters
# session number is cuts column
gen_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")
gen_poster_session_posters_individual$`Session Id` <- gen_poster_session_posters_individual$cuts
# 422 posters in the general session

# grab the two general sessions from the 
# ah posters from the other sheet (very inconvenient!)
ah_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "AFTER_HOURS", skip = 11)
ah_poster_session_posters_individual$`Session Id` <- ah_poster_session_posters_individual$cuts

@@

org_and_after_hours_posters_individual <- 
all_abstracts %>% 
  filter(`Session Type` == "Poster Symposium") %>% 
  filter(Role == "Presenter")
# 196 posters in org + ah sessions

@@

@@

poster_session_durations_no_th_ev
```

@@

```{r}
# to avoid overfilling day-parts
buffer <-7 # is the largest to avoid overfill

@@

# we're going to change these
  all_posters_individual_no_ah_tmp <- all_posters_individual_no_ah
  container_for_this_day_part <- 0
  poster_session_durations_no_th_ev$numbers_of_posters <- 0
  poster_session_durations_no_th_ev$n_posters <- 0
  poster_session_durations_no_th_ev$n_sessions <- 0
  poster_session_durations_no_th_ev$c_sessions <- 0
  poster_session_durations_no_th_ev$c_posters <- 0
  all_of_the_day_parts <- 0
  i <- 1

@@

    # how many rows (posters) to we have remaning to allocate to a day-part?
  print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))

@@

    print(glue('cumulative number of posters: {cumulative_number_of_posters}, j = {j}, i = {i}'))

@@

   # add an identifier so we can see where we have assigned each session so far
 container_for_this_day_part_with_session_from_while <- 
   container_for_this_day_part %>% 
   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$day_part[i]))

@@

 print(poster_session_durations_no_th_ev %>% select(n_max, 
                                                   day_part,
                                                   n_posters,
                                                   n_sessions, 
                                                   c_sessions,
                                                   c_posters))

@@

```{r}
# we have 532 posters not in ah
x1 <- 
all_posters_individual_no_ah

x2 <- 
all_of_the_day_parts

# what sessions are not in there?
session_output <- 
x2 %>% 
  group_by(`Session Id`) %>% 
  pull(`Session Id`) %>% 
  unique

session_input <- 
  x1 %>% 
  group_by(`Session Id`) %>% 
  pull(`Session Id`) %>% 
  unique

setdiff(session_input, session_output)

x1 %>% 
  filter(is.na(Title))

# what are in the output of this loop that is not in the input?
# duplicates?
all_of_the_day_parts %>% 
  distinct()

# why is each poster in each day-part?
all_of_the_day_parts %>% 
 arrange(uid) %>% View

# how many abstracts
posters_with_abstract_ids <- 
all_of_the_day_parts %>% 
  filter(!is.na(`Abstract Id`)) %>% 
  group_by(day_part_from_while) %>% 
  tally() %>% 
  summarise(sum(n)) %>% 
  pull()
```

We have `r posters_with_abstract_ids` posters with Abstract IDs (and so not chair/co-chairs). Compare to `r nrow(gen_and_org_poster_session_posters)` posters that we started with, and space available for `r sum(poster_session_durations$'N max')` posters.

What do we stil need to do?

- Posters

>>>>

# Here we get all the general session posters
# session number is cuts column
gen_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")
gen_poster_session_posters_individual$`Session Id` <- as.character(gen_poster_session_posters_individual$cuts)
# 422 posters in the general session

# In an attempt to solve the scheduling problem...
# Let's make each gen session poster a unique session to see if we can allocat them ok into the slots
# gen_poster_session_posters_individual$`Session Id` <-
#   as.character(glue('9999_{1:nrow(gen_poster_session_posters_individual)}_{gen_poster_session_posters_individual$`Session Id`}'))

# grab the two general sessions from the 
# ah posters from the other sheet (very inconvenient!)
ah_poster_session_posters_individual <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "AFTER_HOURS", skip = 11)
ah_poster_session_posters_individual$`Session Id` <- as.character(ah_poster_session_posters_individual$cuts)

@@

org_and_after_hours_posters_individual <- 
all_abstracts %>% 
  filter(`Session Type` == "Poster Symposium") %>% 
  filter(Role == "Presenter") %>% 
  mutate(`Session Id`  = as.character(`Session Id`))
# 196 posters in org + ah sessions

@@

There are a small number of people with posters who are also discussants, etc. So we need to be sure we don't schedule posters at the same time as those activities.

```{r}
# Do people giving posters have other roles?
# check 
people_with_2_non_poster_roles <- 
all_abstracts %>% 
  group_by(`First Name`, `Last Name`) %>% 
  add_tally() %>% 
  filter(n >= 2) %>% 
  filter(!grepl("Poster", `Session Type`)) %>% 
  mutate(uid = glue('{`First Name`} {`Last Name`}'))

people_in_posters_with_2_roles <- 
  all_abstracts %>% 
  group_by(`First Name`, `Last Name`) %>% 
  add_tally() %>% 
  filter(n >= 2) %>% 
  filter(grepl("Poster", `Session Type`)) %>% 
  mutate(uid = glue('{`First Name`} {`Last Name`}'))

# any overlap?
people_with_posters_and_another_role <- 
people_in_posters_with_2_roles$uid[people_in_posters_with_2_roles$uid %in% people_with_2_non_poster_roles$uid]
```

Here's an experiment to break up the general sessions to see if we can improve the fit into the SAA session limits. We split the general sessions with more than so many papers.

```{r}
gen_poster_session_posters_individual %>%
  group_by(`Session Id`) %>%
  tally() %>%
  ggplot(aes(n)) +
  geom_histogram()

n_allowed <- 8 # almost each poster is it's own session
gen_poster_sessions_n_more_than_n <-
gen_poster_session_posters_individual %>%
  group_by(`Session Id`) %>%
  tally()  %>%
  filter(n >= n_allowed) %>%
  pull(`Session Id`)

gen_posters_in_sessions_n_more_than_n <-
gen_poster_session_posters_individual %>%
  filter(`Session Id` %in% gen_poster_sessions_n_more_than_n) %>%
  group_by(`Session Id`) %>%
  mutate(label = 1:n()) %>%
  mutate(new_session_id1 =  label %% n_allowed) %>%
  group_by(new_session_id1, idx = cumsum(new_session_id1 == 1L)) %>% 
  ungroup() %>%
  mutate(`Session Id` = glue('{`Session Id`}_{idx}'))  %>%
  select(-label, -new_session_id1, -idx )

gen_posters_in_sessions_n_more_than_n %>%
  group_by(`Session Id`) %>%
  tally() %>%
  ggplot(aes(n)) +
  geom_histogram()

# now update the all_posters_individual data frame
all_posters_individual <-
all_posters_individual %>%
  mutate(`Session Id` = as.character(`Session Id`)) %>%
  # first remove the old gen session with more than 12 papers each
  filter(!`Session Id` %in% gen_poster_sessions_n_more_than_n) %>%
  # then add in those same posters, but with their new session ids
  bind_rows(gen_posters_in_sessions_n_more_than_n)
```

@@

# see if we can do the poster day-parts individually

poster_session_durations_no_th_ev <- 
  poster_session_durations %>% 
  filter(day_part != "Th-evening") %>% 
  mutate(poster_day_part = glue('poster-{day_part}')) %>% 
  group_by(poster_day_part) %>% 
  mutate(label = 1:n()) %>% 
  ungroup() %>% 
  mutate(poster_day_part = glue('{poster_day_part}_{label}')) %>% 
  select(-label) %>% 
  mutate(n_max = `N max`)
```

@@

```{r}
# to avoid overfilling day-parts
buffer <- 5 # is the largest to avoid overfill

@@

# we're going to change these
  all_posters_individual_no_ah_tmp <- all_posters_individual_no_ah
  container_for_this_day_part <- 0
  poster_session_durations_no_th_ev$n_posters <- 0
  poster_session_durations_no_th_ev$n_sessions <- 0
  poster_session_durations_no_th_ev$c_sessions <- 0
  poster_session_durations_no_th_ev$c_posters <- 0
  all_of_the_day_parts <- 0
  i <- 1

@@

    # how many rows (posters) to we have remaning to allocate to a day-part?
    # print(glue('number of poster sessions remaining to classify is {nrow(all_posters_individual_no_ah_tmp)}'))

@@

    # print(glue('cumulative number of posters: {cumulative_number_of_posters}, j = {j}, i = {i}'))

@@

   # add an identifier so we can see where we have assigned each session so far
 container_for_this_day_part_with_session_from_while <- 
   container_for_this_day_part %>% 
   mutate(day_part_from_while  = as.character(poster_session_durations_no_th_ev$poster_day_part[i]))

@@

 print(poster_session_durations_no_th_ev %>% select(n_max, 
                                                   poster_day_part,
                                                   n_posters,
                                                   n_sessions, 
                                                   c_sessions,
                                                   c_posters))

@@

```{r}

```

We have no clash with paper presenters.

What do we still need to do?
